Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-02-10 Paper/Manuscript.tex
Compared paper and refactored code for the single target until the DFT.

\subsection{Power and angle samples}
The transceiver measures voltage over a load, $\zrxe$, connected in series with the transducer impedance, $\ztde$. When calculating various acoustic properties a system gain parameter will be used which assumes a matched receiver load. The total received power, $\prxe(\samplesymt)$, from all transducer sectors for a matched receiver load (Fig. \ref{fi:impedances}) is given by: 
\begin{equation}
\label{eq:prx}
\prxe(\samplesymt) = \nchannels\left( \frac{|\ypc(\samplesymt)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{\zrxe} \right)^2 \frac{1}{|\ztde|}.
\end{equation}
%
\begin{figure}
\includegraphics[width=\reprintcolumnwidth]{Fig_impedances}
\caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
\end{figure}

Forward/aft and port/starboard phase angles of target echoes are estimated by combining the transducer half signals thus: 
%
\begin{eqnarray}
\label{eq:phase1}
y_\along(\samplesymt) & = & y_{\textrm{pc,fore}}(\samplesymt) y_{\textrm{pc,aft}}^*(\samplesymt), \\
y_\athw(\samplesymt) & = & y_{\textrm{pc,star}}(\samplesymt) y_{\textrm{pc,port}}^*(\samplesymt),
\end{eqnarray}
%
where $y_\along(\samplesymt)$ is the electrical angle along the minor axis of the transducer (positive in the forward direction when ship-mounted) and $y_\athw(\samplesymt)$ the electrical angle along the major axis of the transducer (positive to starboard when ship-mounted), where complex signals are represented in the form $e^{j 2\pi \freqsym \timesym}$, where $j = \sqrt{-1}$. The physical echo arrival angles ($\along$ and $\athw$) are then given by:
%
\begin{eqnarray}
\label{eq:phase2}
\along(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\along(\samplesymt)), \Re(y_\along(\samplesymt) \right)}{\anglefalong}\right) \\
\athw(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\athw(\samplesymt)), \Re(y_\athw(\samplesymt) \right)}{\anglefathw}\right).
\end{eqnarray}
%
where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles and are derived from the transducer geometry and $\fc$ the centre frequency of the chirp pulse \citep{ehrenberg1979}. The inverse sine is indicated by $\arcsin$, the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$. As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for port/starboard angles.
%
\begin{figure}
\includegraphics[width=\reprintcolumnwidth]{'Fig_theta_phi.png'}
\caption{\label{fi:theta_phi}Physical angles.}
\end{figure}

@@

Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated. This is typically achieved through a single echo detection algorithm (SED). Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets. The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} $\ypc(\samplesymt)$ is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively. A simple pseudo SED algorithm is implemented in the code for illustrative purposes.

@@

The discrete Fourier transforms of the target signal, $\ypctargetf(\samplesymf)$, and the reduced auto correlation signal, $\ymfautoredf(\samplesymf)$, are given by:
\begin{eqnarray}
\label{eq:DFT_Target_Auto}
\ypctargetf(\samplesymf) & = & \dft_\ndft(\ypctarget(\samplesymt)),\\
\ymfautoredf(\samplesymf) & = & \dft_\ndft(\ymfautored(\samplesymt)),
\end{eqnarray}
where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, is then calculated by: 
%
\begin{equation}
\label{eq:DFT_Target_Auto_Norm}
\ypctargetnormf(\samplesymf) = \frac{\ypctargetf(\samplesymf)} {\ymfautoredf(\samplesymf)}.
\end{equation}
%
Assuming, as a first approximation, that the impedances of the transceiver and transducer are independent of frequency, the received power into a matched load, $\prxetf(\samplesymf)$, is then estimated by:
\begin{equation}
\label{eq:prx_FFT_target}
\prxetf(\samplesymf) = \nchannels\left( \frac{|\ypctargetnormf(\samplesymf)|}{2\sqrt{2}} \right)^2 
\left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \frac{1}{|\ztde|}, % All channels (*4), matched load (/2), and effective values (/sqrt(2))
\end{equation}
%
noting that any variation of impedance with frequency will be reflected in the $\gainzero$ obtained from the calibration process.

>>>>

\subsection{Power and angle samples}
The transceiver measures voltage over a load, $\zrxe$, connected in series with the transducer impedance, $\ztde$. When calculating various acoustic properties a system gain parameter will be used which assumes a matched receiver load. The total received power, $\prxe(\samplesymt)$, from all transducer sectors for a matched receiver load (Fig. \ref{fi:impedances}) is given by: 
\begin{equation}
\label{eq:prx}
\prxe(\samplesymt) = \nchannels\left( \frac{|\ypc(\samplesymt)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{\zrxe} \right)^2 \frac{1}{|\ztde|}.
\end{equation}
%
\begin{figure}
\includegraphics[width=16cm]{Fig_impedances}
\caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
\end{figure}

Forward/aft and port/starboard phase angles of target echoes are estimated by combining the transducer half signals thus: 
%
\begin{eqnarray}
\label{eq:phase1}
y_\along(\samplesymt) & = & y_{\textrm{pc,fore}}(\samplesymt) y_{\textrm{pc,aft}}^*(\samplesymt), \\
y_\athw(\samplesymt) & = & y_{\textrm{pc,star}}(\samplesymt) y_{\textrm{pc,port}}^*(\samplesymt),
\end{eqnarray}
%
where $y_\along(\samplesymt)$ is the electrical angle along the minor axis of the transducer (positive in the forward direction when ship-mounted) and $y_\athw(\samplesymt)$ the electrical angle along the major axis of the transducer (positive to starboard when ship-mounted), where complex signals are represented in the form $e^{j 2\pi \freqsym \timesym}$, where $j = \sqrt{-1}$. The physical echo arrival angles ($\along$ and $\athw$) are then given by:
%
\begin{eqnarray}
\label{eq:phase2}
\along(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\along(\samplesymt)), \Re(y_\along(\samplesymt) \right)}{\anglefalong}\right) \\
\athw(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\athw(\samplesymt)), \Re(y_\athw(\samplesymt) \right)}{\anglefathw}\right).
\end{eqnarray}
%
where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles and are derived from the transducer geometry and $\fc$ the centre frequency of the chirp pulse \citep{ehrenberg1979}. The inverse sine is indicated by $\arcsin$, the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$. As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for port/starboard angles.
%
\begin{figure}
\includegraphics[width=16cm]{Fig_theta_phi}
\caption{\label{fi:theta_phi}Physical angles.}
\end{figure}

@@

Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated. This is typically achieved through a single echo detection algorithm (SED). Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets. The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} $\ypc(\samplesymt)$ (or peak $S_p(n)$?? current code uses $\ypc$) is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively. A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
%
\begin{figure}
%\includegraphics[width=\reprintcolumnwidth]{'Fig_SED'}
\caption{\label{fi:SED}Illustrate the single target data somehow.}
\end{figure}

@@

The discrete Fourier transforms of the target signal, $\ypctargetf(\samplesymf)$, and the reduced auto correlation signal, $\ymfautoredf(\samplesymf)$, are given by:
\begin{eqnarray}
\label{eq:DFT_Target_Auto}
\ypctargetf(\samplesymf) & = & \dft_\ndft(\ypctarget(\samplesymt)),\\
\ymfautoredf(\samplesymf) & = & \dft_\ndft(\ymfautored(\samplesymt)),
\end{eqnarray}
where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (\ref{fi:SED}) is then calculated by: 
%
\begin{equation}
\label{eq:DFT_Target_Auto_Norm}
\ypctargetnormf(\samplesymf) = \frac{\ypctargetf(\samplesymf)} {\ymfautoredf(\samplesymf)}.
\end{equation}
%
The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...
%
\begin{figure}
\includegraphics[width=16cm]{Fig_normalizedDFT}
\caption{\label{fi:SED}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
\end{figure}

Assuming, as a first approximation, that the impedances of the transceiver and transducer are independent of frequency, the received power into a matched load, $\prxetf(\samplesymf)$, is then estimated by:
\begin{equation}
\label{eq:prx_FFT_target}
\prxetf(\samplesymf) = \nchannels\left( \frac{|\ypctargetnormf(\samplesymf)|}{2\sqrt{2}} \right)^2 
\left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \frac{1}{|\ztde|}, % All channels (*4), matched load (/2), and effective values (/sqrt(2))
\end{equation}
%
noting that any variation of impedance with frequency will be reflected in the $\gainzero$ obtained from the calibration process.
